Effects of gluon kinematics and the Sudakov\\ form factor on the dipole amplitude

### 2023-05-19 Paper/draft.tex
paper update

\title{Effects of gluon kinematics and the Sudakov form factor\\ on the dipole amplitude}

@@

\begin{abstract}
We investigate effects of exact gluon kinematics on the parameters of the
Golec-Biernat--W\"usthoff, and Bartels--Golec-Biernat--Kowalski saturation
models. The resulting parameters show some differences, particularly, the
normalization of the dipole cross section $\sigma_0$. The refitted models are
used for the dijet production process in DIS to investigate effects of the
Sudakov form factor at Electron Ion Collider energies.  
\end{abstract}

@@

\comment{preprint number (SS)}\\

\begin{figure}[t]
\centering
\resizebox{0.6\textwidth}{0.2\textwidth}{
\begin{tikzpicture}
  \draw[snake=coil,segment aspect=0,segment length=0.1in](-0.875in , 0.5in)--(-0.5in,0.in);
  \draw[snake=coil,segment aspect=0,segment length=0.1in](0.875in ,0.5in)--(0.5in,0.in);
  \draw(-0.5in , 0in)--(0.5in,0in)--(0.5in , -0.5in)--(-0.5in,-0.5in)--cycle ;

@@

  \draw[ snake=coil,segment aspect=1.5,segment length=0.1in](0.5in , -1in)--(0.5in,-0.5in);
  \draw[snake=coil,segment aspect=1.5,segment length=0.1in](-0.5in , -1in)--(-0.5in,-0.5in);
\end{tikzpicture}
}
\caption{Kinematic variables of the gluon contribution to $F_2$. The momentum fraction of gluon is $x/z$ as it has to contribute via the quark box.
}
\label{fig:gluon-kinem}
\end{figure}

@@

One of the main type of processes that is suited particularly well to the studies
of proton structure is the {\it deep inelastic
scattering}~(DIS)~\cite{Roberts:1990ww,Collins:2011zzd}. The data from HERA has
enabled us to answer many questions in that domain and largely improved our
picture of interior of a proton. 
It is therefore very exiting that the next generation of a DIS machines,
the election-ion collider~(EIC)~\cite{NAP25171} is making its way and will soon
allow us to uncover more details about the hadron structure.

With an increase of the of the center of mass energy at fixed value of the
photon virtuality, $Q$, one shall probe  the region of the small Bjorken scaling
variable, $x$.  
In that region, dominated by gluons\cite{Roberts:1990ww,Collins:2011zzd,
Kovchegov:2012mbw}, the appropriate theoretical formalism to use is the Color
Glass Condensate~\cite{McLerran:1993ni, Ayala:1995kg, Kovchegov:1996ty,Jalilian-Marian:1996mkd,Iancu:2000hn,Iancu:2001ad,Ferreiro:2001qy,Iancu:2002xk} or, if the density of gluons is still moderate, the
High Energy Factorization also called $k_T$-factorization~\cite{Catani:1990xk,Catani:1990eg,Collins:1991ty,Catani:1993ww,Catani:1993rn,Catani:1994sq}.

In the above frameworks, the interaction of the photon with gluons happens
through a box of quarks, as depicted in Fig.~\ref{fig:gluon-kinem}. An object,
which is central to the description of this process is called the dipole gluon
density, $\fdp(x,k_T^2)$, which is a type of {\it transverse momentum
dependent} PDFs (TMD
PDFs)~\cite{Bomhof:2006dp,Dominguez:2010xd,Dominguez:2011wm,Xiao:2017yya}. 
%
It appears in the $k_T$-factorization
formula~\cite{Catani:1990eg,Dominguez:2010xd,Dominguez:2011wm,Xiao:2017yya},
which we shall discuss in the next section, and plays a role similar to that of
the collinear parton distribution function in the collinear factorization.

Much effort has been made
to study this object, notably the work of Balitsky, Fadin, Kuraev and Lipatov
(BFKL)~\cite{Balitsky:1978ic, Kuraev:1977fs} has predicted sharp rise of cross
sections with $1/x$.  However, such rise violates unitarity and the Froissart
bound~\cite{Roberts:1990ww,Collins:2011zzd,Kovchegov:2012mbw,Barone:1993sy}. In
\cite{Gribov:1983ivg} it has been recognized that gluon recombination can tame
the growth of gluons \cite{Gribov:1983ivg}. The interplay of linear and
nonlinear terms leads to a phenomenon called {\it gluon saturation} and
corresponding evolution equations are known as
Balitsky-Kovchegov~(BK)~\cite{Balitsky:1995ub,Kovchegov:1999yj} equation or
JIMWLK\cite{Kovner:1999bj,Kovner:2000pt,Iancu:2000hn,JalilianMarian:1997dw,JalilianMarian:1997gr,JalilianMarian:1997jx} equation.

Due to its clear picture, one often uses a position space counter part of the gluon density, called {\it dipole cross section}. Despite the success of the aforementioned evolution equations, there exists a number of phenomenological models of the dipole cross section, which are popular due to to their simplicity. 
\commentPending{More about saturation models}
A notable example of a dipole saturation model is the one proposed by
Golec-Biernat and Wu\"sthoff (GBW)~\cite{Golec-Biernat:1998zce} and the one proposed by
Bartels, Golec-Biernat and Kowalski (BGK)~\cite{Bartels:2002cj}. We will use these two models in this study. There are other models, with different considerations, such as impact parameter dependence~({\it cf.}
\cite{Kowalski:2003hm,McLerran:1993ni,Forshaw:1999uf,Iancu:2003ge}).

\commentPending{Section about what we did: fits of saturation models with exact
kinematics, use of those results + Sudakov for EIC predictions.}

In the present study, we fit the GBW~\cite{Golec-Biernat:1998zce} and BGK~\cite{Bartels:2002cj} models in the $k_T$ factorization formula of $F_2$, instead of the original dipole factorization formula. As discussed below, this allows us to fit the model with exact gluon kinematics. 
Since inclusive observables do not reveal the $k_T$ dependence well, we also consider a less inclusive observable which is more sensitive to the $k_T$ dependence of the gluon distribution.
The resulting models are put to the test by applying them to the predictions of dijet DIS at EIC, including effects of the Sudakov form factor, which resums logarithms of the small transverse momentum.

The paper is organized as follows.  In the next section, we present theoretical
framework which we used and, in particular, outline differences of the
$k_T$-factorization and the dipole-factorization.  In
Sec.~\ref{sec:F2fits}, results of the new fits to HERA data in the
$k_T$-factorization formula are presented. 
In Sec.~\ref{sec:dijetsEIC}, we apply the results from the previous section to
compute distributions for dijet production in DIS at the Electron-Ion Collider
(EIC)~\cite{NAP25171} and make comparisons to our earlier study of Ref.~\cite{vanHameren:2021sqc}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{The framework}
\commentPending{Missing a general factorization formula}
In general, a factorization formula has the form
\begin{equation}
    \sigma=H \otimes \phi,
\end{equation}
where $\otimes$ denotes convolution. In this form one achieves two things. Firstly, one can confine long distance effects in one of the factors (let $\phi$ be such an object), and leaves $H$ as an perturbatively calculable object. Secondly, one may factorize out a (to some extent) process independent object.  
In the present study, the dipole cross section and the gluon densities both  contain nonperturbative part and are, to an extent, process independent~\footnote{They are not universal, however they may appear in different processes. (see for example Refs.~\cite{Dominguez:2011wm,Kotko:2015ura})}.
In this section, we will outline two factorizations; the $k_T$ factorization and the dipole factorization. \\
As it was mentioned earlier, we use the GBW and BGK models.  
The models were originally formulated in the position-space version of the
$k_T$-factorization formula. %, and the object which corresponds to the gluon
%density of the $k_T$-factorization is called the dipole cross section. 
In the GBW model, the dipole cross section has the form~\cite{Golec-Biernat:1998zce}
\begin{align}
\label{eq:gbw-dipole}
\sigma_{\mathrm{GBW}}(x,r)&=\sigma_0\left(1-e^{-r^2/R^2_0}\right)&\mathrm{where}&
&R^{-2}_0&=\frac{Q_0^{2}}{4}\left(\frac{x_0}{x}\right)^\lambda.
\end{align}
%
The dipole cross section is related to the gluon density by the Fourier
transform
%
\begin{equation}
\alpha_s\fdp(x,k_T^2)=\frac{N_c}{4\pi}\int\frac{d^2\mathbf{r}}{(2\pi)^2}e^{i\mathbf{k}_t\cdot \mathbf{r} }\nabla_{\mathbf{r}}^2\sigma_{\mathrm{dipole}}(x,r).
\label{eq:dipole-gluon}
\end{equation}
%
The essence of the GBW model is encoded in the $x$-dependent saturation scale
$Q_s^2(x)=Q^2_0(x_0/x)^{-\lambda}$, which separates the saturation region and
the scaling region.  A extension of the above model was proposed by Bartels, Golec-Biernat
and Kowalski~(BGK)~\cite{Bartels:2002cj} who incorporated the DGLAP evolution
in the GBW cross section~(\ref{eq:gbw-dipole}) by modifying the exponent to 
%
\begin{equation}
R_0^{-2}=\frac{\pi^2\alpha(\mu^2)xg(x,\mu^2)}{3\sigma_0},
\end{equation}
where
\begin{align}
\mu^2&=\frac{C}{r^2}+\mu_0^2 & &\mathrm{and} & xg(x,Q^2_0)&=A_g x^{-\lambda_g}(1-x)^{5.6},
\end{align}
%
thus improving the description of data at higher $Q^2$.\\

@@

Firstly, let us outline the factorization formulas.
With $q'\equiv q+x p$, one decomposes $k$ and $\kappa$, defined in
Fig.~\ref{fig:gluon-kinem}, as
%
\begin{align}
    \kappa&=\alpha p-\beta q'+\kappa_t&&\mathrm{and}& k&=a p- bq'+k_T.
\end{align}
For the contribution depicted in Fig.~\ref{fig:gluon-kinem}, the structure function, $F_2$ for transversely and longitudinally polarized photons, factorize, with a change of variable  ${\boldsymbol{\kappa}'}_t\equiv{\boldsymbol{\kappa}_t}-(1-\beta)\mathbf{k}_t$, in the form~\cite{ Kimber:2001uaa,Kwiecinski:1997ee}
\begin{multline}
	F_2(x,Q^2)=\sum_f e_f^2 \frac{Q^2}{2\pi}\int\frac{dk^2_t}{k_T^2}\int^1_0d\beta\int d{\kappa'}_t\alpha_s(\mu^2) \fdp(x/z,k_T^2)\Theta(1-x/z)\\
	\left[\left(\beta^2+(1-\beta)^2\right)\left(\frac{I_1}{2\pi}-\frac{I_2}{2\pi}\right)
	+\left(m_f^2+4Q^2\beta^2(1-\beta)^2\right)\left(\frac{I_3}{2\pi}-\frac{I_4}{2\pi}\right)\right],
	\label{eq:angle-integrated}
\end{multline}
where
\begin{equation}
	\frac{1}{z}=1+\frac{{\kappa'}^2_t+m_f^2}{\beta(1-\beta)Q^2}+\frac{k^2_t}{Q^2}\,.
	\label{eq:z}
\end{equation}

One should note, that gluons are not probed directly, thus the argument of the
gluon density is $x/z$ rather than $x$.  If one, instead, uses $ 1/z=1+4
m_f^2/Q^2$ and assumes that $\mu$ is independent of $\kappa'$ and $k_T$, the
above formula can be written in the impact parameter
space~\cite{Golec-Biernat:1998zce,Nikolaev:1990ja} ({\it i.e.} as a
dipole-factorization formula)
%
\begin{equation}
  F_2\left(x,Q^2\right)=\sum_f e_f^2 \frac{Q^2}{4\pi^2} \int^1_0d\beta \int \frac{d^2\mathbf{r}}{(2\pi)^2} \left|\Psi\left(\tilde{x},\beta,Q^2\right)\right|^2\sigma_{\mathrm{dipole}}\left(\tilde{x},r\right),
  \label{eq:dipole-factorization}
\end{equation}

where the photon wave function, $\left|\Psi\left(x,\beta,Q^2\right)\right|^2$,
describes splitting of the incoming photon into a $q\overline{q}$ pair with
light-cone momenta fractions $\beta$ and $1-\beta$ respectively, and the dipole
cross section, $\sigma_{\mathrm{dipole}}\left(\tilde x,r\right)$, describes the
interaction of the colour dipole of the size $r$ with the proton. In
Ref.~\cite{Golec-Biernat:1998zce}, $\tilde{x}=x(1+4m_f/Q^2)$ with nonzero value
of $m_f$ was used even for light flavours, which is necessary to partially
substitute $k_T$ and ${\kappa'}_T$ in Eq.~(\ref{eq:z}).
%The original GBW and BGK models were fitted in the dipole factorization, Eq.~(\ref{eq:dipole-factorization}), while 
In the present study, we use Eq.~(\ref{eq:angle-integrated}) to fit the models, where the gluon density is obtained by evaluating Eq.~(\ref{eq:dipole-gluon}).
%Comparing, the GBW and BGK models, one identifies that, the GBW model is effectively the BGK model with 
%\begin{equation}
%\begin{split}
%    \alpha_s(\mu^2)xg(x,\mu^2)\rightarrow &\alpha_s(Q_0^2)xg(x,Q_0^2)\\
%    &=\alpha_s x^{-\lambda}(1-x)^{5.6}.
%\end{split}
%\end{equation}
\commentPending{??}{\color{blue}response: the argument on why we consider $\alpha$ as a constant is removed, as it is not a solid argument.}
%For this reason,
In order to account for the running coupling in Eq.~(\ref{eq:angle-integrated}), 
we assume that $\alpha_s$ in Eq.~(\ref{eq:dipole-gluon}) is
constant for the GBW model, and thus explicitly multiply by the running coupling
%
\begin{equation}
\alpha_s(\mu^2)\fdp(x,k_T^2)\rightarrow \alpha_s(\mu)\frac{\alpha_s\fdp(x,k_T^2)}{0.2},
\end{equation}
where
%
\begin{equation}
\alpha_s(\mu^2)=\frac{1}{\frac{11 C_A-2n_f}{12\pi}\log\left(\frac{\mu^2}{\Lambda_{\mathrm{QCD}}^2 }\right)},
\end{equation}
and  $\Lambda_{\mathrm{QCD}}^2=0.09\;\GeV^2$.

The factor 0.2 is an arbitrary normalization whose effect is absorbed by $\sigma_0$ and hence, bears no importance.
(For a more detailed analysis of the dipole gluon density from the BGK model see Ref.~\cite{Luszczak:2022fkf}\footnote{In Ref.~\cite{Luszczak:2022fkf}, the coupling constant was treated differently, and the large $k_T$-region was matched to the derivative of $xg(x)$, while we have used series transformation to accelerate the high-$k_T$ integration. Their treatment is also different from the original BGK paper~\cite{Bartels:2002cj}. As a consequence their gluon density is positive in the large $k_T$-region and the overall large-$k_T$ behaviour is significantly different.}
)
While there is some ambiguity on what the argument of $\alpha_s(\mu^2)$ should be, we follow Ref.~\cite{Kwiecinski:1997ee} and use
\begin{equation}
	\mu^2=k_T^2+{\kappa'}_t^2+m_f^2+\mu^2_0,
\end{equation}
where we add $\mu^2_0=4\;\GeV^2$ in order to freeze the coupling at low scales.

@@

We fitted the GBW and BGK models to the $F_2$ data from
HERA~\cite{Abt:2017nkc}. A numerical program was written with help of
CERNLIB~(DPSIPG)~\cite{Kolbig:1972zz}, GSL~\cite{GSL}, CUBA~\cite{Hahn:2004fe}
and ROOT~\cite{Brun:1997pa} libraries to evaluate $F_2$. The data were selected
to be in the range $0.045\leq Q^2\leq 650\;\mathrm{GeV^2}$, $x<0.01$.  As with
the previous fit~\cite{Goda:2022wsc}, the $c$ and $b$ flavours were taken into
account with the mass $1.3$ and $4.6\;\mathrm{GeV}$, respectively. As discussed
earlier, we take the light quarks to be massless.  The fitting was performed
using MnMigrad and MnSimplex of ROOT::Minuit2~\cite{James:2004xla}.

We studied the following cases
%
\begin{itemize}
\item GBW model with the fixed coupling in $k_T$-factorization  ($k_T$-GBW),
\item GBW model with the running coupling in $k_T$-factorization (rc-$k_T$-GBW),
\item BGK model in $k_T$-factorization ($k_T$-BGK),
\end{itemize} 
%
and, as a reference, we provide the following results from
Ref.~\cite{Goda:2022wsc};
%
\begin{itemize}
\item GBW model with massless light quarks in the dipole-factorization ($r$-GBW),
\item GBW model with massive light quarks in the dipole-factorization ($r$-GBW-massive),
\item  BGK model in  the dipole-factorization ($r$-BGK).
\end{itemize}

>>>>

\title{
Effects of gluon kinematics and the Sudakov\\
form factor on the dipole amplitude
}

@@

\begin{abstract}
We investigate effects of exact gluon kinematics on the parameters of the
Golec-Biernat--W\"usthoff, and Bartels--Golec-Biernat--Kowalski saturation
models. The resulting parameters show some differences, particularly, in the
normalization of the dipole cross section $\sigma_0$. The refitted models are
used for the dijet production process in DIS to investigate effects of the
Sudakov form factor at Electron Ion Collider energies.  
\end{abstract}

@@

\comment{preprint number (SS)}\\
\comment{the information on energy and cuts is missing on the plots (KK)}
\comment{I reformulated some sentences added references and  interpretation of the curves where we use the KS gluon.  }

\begin{figure}[t]
\centering
\resizebox{0.6\textwidth}{0.2\textwidth}{
\begin{tikzpicture}
  \draw(-0.75in,0.25in)node[left,scale=0.02in]{$q$};
  \draw[snake=coil,segment aspect=0,segment length=0.1in](-0.875in , 0.5in)--(-0.5in,0.in);
  \draw[snake=coil,segment aspect=0,segment length=0.1in](0.875in ,0.5in)--(0.5in,0.in);
  \draw(-0.5in , 0in)--(0.5in,0in)--(0.5in , -0.5in)--(-0.5in,-0.5in)--cycle ;

@@

  \draw[ snake=coil,segment aspect=1.5,segment length=0.1in](0.5in , -1in)--(0.5in,-0.5in);
  \draw[snake=coil,segment aspect=1.5,segment length=0.1in](-0.5in , -1in)--(-0.5in,-0.5in);
\end{tikzpicture}
}
\caption{Kinematic variables of the inclusive DIS. \commentPending{Photon momentum (TG)}}
\label{fig:gluon-kinem}
\end{figure}

@@

One of the main type of processes that is suited particularly well to the
studies of proton structure is the {\it deep inelastic
scattering}~(DIS)~\cite{Roberts:1990ww,Collins:2011zzd}. The data from HERA has
enabled us to answer many questions in that domain and largely improved our
picture of interior of a proton.  It is therefore very exiting that the next
generation of a DIS machines, the Election Ion Collider~(EIC)~\cite{NAP25171} is
making its way and will soon allow us to uncover more details about the hadron
structure.

At large center of mass energy and fixed value of the photon virtuality, $Q$,
one probes  the region of small Bjorken scaling variable, $x$.  In that region,
dominated by gluons\cite{Roberts:1990ww,Collins:2011zzd, Kovchegov:2012mbw}, the
approach of High Energy Factorization, also called
$k_T$-factorization~\cite{Catani:1990xk,Catani:1990eg,Collins:1991ty,Catani:1993ww,Catani:1993rn,Catani:1994sq},
has proven to be suited particularly well.  In this framework, the interaction
of the photon with gluons happens through a box of quarks, as depicted in
Fig.~\ref{fig:gluon-kinem}. An object, which is central to the description of
this process is called the dipole gluon density, $\fdp(x,k_T^2)$, which is a
type of {\it transverse momentum dependent} PDFs (TMD
PDFs)~\cite{Bomhof:2006dp,Dominguez:2010xd,Dominguez:2011wm,Xiao:2017yya}. 
%
Due to its clear picture, one often uses a position-space counterpart of the gluon density, called {\it dipole cross section}. 
%
It appears in the $k_T$-factorization
formula~\cite{Catani:1990eg,Dominguez:2010xd,Dominguez:2011wm,Xiao:2017yya},
which we shall discuss in the next section, and plays a role similar to that of
the collinear parton distribution function in the collinear factorization.

Much effort has been made to study this object, notably the work of Balitsky,
Fadin, Kuraev and Lipatov (BFKL)~\cite{Balitsky:1978ic, Kuraev:1977fs} has
predicted sharp rise of cross sections with $1/x$.  However, such rise violates
unitarity and the Froissart
bound~\cite{Roberts:1990ww,Collins:2011zzd,Kovchegov:2012mbw,Barone:1993sy}. In
Ref.~\cite{Gribov:1983ivg} it has been recognized that gluon recombination can
tame the growth of gluons. The interplay of linear and
nonlinear terms leads to a phenomenon called {\it gluon saturation} and
corresponding evolution equations are known as
Balitsky-Kovchegov~(BK)~\cite{Balitsky:1995ub,Kovchegov:1999yj} equation or
JIMWLK\cite{Kovner:1999bj,Kovner:2000pt,Iancu:2000hn,JalilianMarian:1997dw,JalilianMarian:1997gr,JalilianMarian:1997jx}
equation.

Despite the success of the aforementioned evolution equations, there exist a
number of phenomenological models of the dipole cross section, which are popular
due to to their simplicity.  A notable example is the model proposed by
Golec-Biernat and Wu\"sthoff (GBW)~\cite{Golec-Biernat:1998zce} and its
extension by Bartels, Golec-Biernat and Kowalski (BGK)~\cite{Bartels:2002cj}. We
will use these two specific models in our study. We note, however, that other
saturation models have been discussed in the
literature~\cite{Kowalski:2003hm,McLerran:1993ni,Forshaw:1999uf,Iancu:2003ge}.

In the present study, we fit the GBW and BGK models to HERA
data~\cite{Abt:2017nkc} in the $k_T$-factorization formula of $F_2$, instead of
the original dipole factorization formula. As discussed below, this allows us to relatively easy investigate the role 
of the  exact gluon kinematics.  Since inclusive observables do not
reveal the $k_T$ dependence well, in order to demonstrate the effects of exact kinematics we consider a less inclusive observable
which is more sensitive to the $k_T$ dependence of the gluon distribution.  The
resulting models are used for predictions of dijet production in DIS at EIC,
including effects of the Sudakov form factor, which resums logarithms of the
small transverse momentum. 
%
This process is known to be sensitive to another type of gluon density, called
Weitzs\"acker-Williams~\cite{Dominguez:2010xd,Dominguez:2011wm}. For that
process, we study, in particular, the angular correlation of jets, and that of
the scattered electron and the jets.

The paper is organized as follows.  In the next section, we present theoretical
framework which we use and outline differences between the $k_T$-factorization
and the dipole factorization.  In Sec.~\ref{sec:F2fits}, results of the new fits
to HERA data in the $k_T$-factorization formula are presented.  In
Sec.~\ref{sec:dijetsEIC}, we apply the results from the previous section to
compute distributions for dijet production in DIS at the EIC and make
comparisons to our earlier study of Ref.~\cite{vanHameren:2021sqc}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{The framework}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The DIS cross section (structure function) factorises and can be written in a
form
%
\begin{equation}
   d\sigma = \sum_{a}  \phi_{a/h} \otimes H_{\gamma a \to X} ,
\end{equation}
%
where the hard function, $H_{\gamma a \to X}$, involving a parton $a$ in the
initial state, is calculable perturbatively, and $\phi_{a/h}$ is the parton
distribution function of $a$ in a hadron $h$.  
%
The symbol $\otimes$ denotes appropriate convolution. All nonperturbative
effects are absorbed in $\phi_{a/h}$, while $H_{\gamma a \to X}$ can be computed
order by order in $\alpha_s$.

We will study factorization in two versions: defined in momentum and position
space respectively.  As mentioned earlier, in the fit we use the GBW and BGK models.  The models were
originally formulated in the position-space version of the $k_T$-factorization
formula.  In the GBW model, the dipole cross section has the
form~\cite{Golec-Biernat:1998zce}
%
\begin{align}
\label{eq:gbw-dipole}
\sigma_{\mathrm{GBW}}(x,r)&=\sigma_0\left(1-e^{-r^2/R^2_0}\right)&\mathrm{where}&
&R^{-2}_0&=\frac{Q_0^{2}}{4}\left(\frac{x_0}{x}\right)^\lambda.
\end{align}
%
The dipole cross section is related to the dipole gluon density by the Fourier
transform
%
\begin{equation}
\alpha_s\fdp(x,k_T^2)=\frac{N_c}{4\pi}\int\frac{d^2\mathbf{r}}{(2\pi)^2}e^{i\mathbf{k}_t\cdot \mathbf{r} }\nabla_{\mathbf{r}}^2\sigma_{\mathrm{dipole}}(x,r).
\label{eq:dipole gluon}
\end{equation}
%
The essence of the GBW model is encoded in the $x$-dependent saturation scale
$Q_s^2(x)=Q^2_0(x_0/x)^{-\lambda}$, which separates the saturation region and
the scaling region.  An extension of the above model was proposed by Bartels, Golec-Biernat
and Kowalski~\cite{Bartels:2002cj} who incorporated the DGLAP evolution
in the GBW dipole cross section~(\ref{eq:gbw-dipole}) by modifying the exponent
to 
%
\begin{equation}
R_0^{-2}=\frac{\pi^2\alpha(\mu^2)xg(x,\mu^2)}{3\sigma_0},
\end{equation}
where
\begin{align}
\mu^2&=\frac{C}{r^2}+\mu_0^2 & &\mathrm{and} & xg(x,Q^2_0)&=A_g x^{-\lambda_g}(1-x)^{5.6},
\end{align}
%
thus improving the description of data at higher $Q^2$.\\

@@

Firstly, let us outline the factorization formulas.
With $q'\equiv q+x p$, one decomposes $k$ and $\kappa$, defined in
Fig.~\ref{fig:gluon-kinem}, as
%
\begin{align}
    \kappa&=\alpha p-\beta q'+\kappa_t&&\mathrm{and}& k&=a p- bq'+k_T.
\end{align}
%
For the contribution depicted in Fig.~\ref{fig:gluon-kinem}, the structure
function $F_2$ factorizes with a change of variable
${\boldsymbol{\kappa}'}_t\equiv{\boldsymbol{\kappa}_t}-(1-\beta)\mathbf{k}_t$,
to the form~\cite{ Kimber:2001uaa,Kwiecinski:1997ee}
%
\begin{multline}
	F_2(x,Q^2)=\sum_f e_f^2 \frac{Q^2}{2\pi}\int\frac{dk^2_t}{k_T^2}\int^1_0d\beta\int d{\kappa'}_t\alpha_s(\mu^2) \fdp(x/z,k_T^2)\Theta(1-x/z)\\
	\left[\left(\beta^2+(1-\beta)^2\right)\left(\frac{I_1}{2\pi}-\frac{I_2}{2\pi}\right)
	+\left(m_f^2+4Q^2\beta^2(1-\beta)^2\right)\left(\frac{I_3}{2\pi}-\frac{I_4}{2\pi}\right)\right],
	\label{eq:angle-integrated}
\end{multline}
where
\begin{equation}
	\frac{1}{z}=1+\frac{{\kappa'}^2_t+m_f^2}{\beta(1-\beta)Q^2}+\frac{k^2_t}{Q^2}\,.
	\label{eq:z}
\end{equation}

One should note that the gluons are not probed directly, thus the argument of
the gluon density is $x/z$ rather than $x$.  If one, instead, uses $ 1/z=1+4
m_f^2/Q^2$ and assumes that $\mu$ is independent of $\kappa'$ and $k_T$, the
above formula can be written in the impact parameter
space~\cite{Golec-Biernat:1998zce,Nikolaev:1990ja} ({\it i.e.} as a dipole
factorization formula)
%
\begin{equation}
  F_2\left(x,Q^2\right)=\sum_f e_f^2 \frac{Q^2}{4\pi^2} \int^1_0d\beta \int \frac{d^2\mathbf{r}}{(2\pi)^2} \left|\Psi\left(\tilde{x},\beta,Q^2\right)\right|^2\sigma_{\mathrm{dipole}}\left(\tilde{x},r\right),
  \label{eq:dipole factorization}
\end{equation}
%
where the photon wave function, $\left|\Psi\left(x,\beta,Q^2\right)\right|^2$,
describes splitting of the incoming photon into a $q\overline{q}$ pair with
light-cone momenta fractions $\beta$ and $1-\beta$ respectively, and the dipole
cross section, $\sigma_{\mathrm{dipole}}\left(\tilde x,r\right)$, describes the
interaction of the colour dipole of size $r$ with the proton. In
Ref.~\cite{Golec-Biernat:1998zce}, $\tilde{x}=x(1+4m_f/Q^2)$ with nonzero value
of $m_f$ was used even for light flavours, which is necessary to partially
substitute $k_T$ and ${\kappa'}_T$ in Eq.~(\ref{eq:z}).
%The original GBW and BGK models were fitted in the dipole factorization, Eq.~(\ref{eq:dipole factorization}), while 
In the present study, we use Eq.~(\ref{eq:angle-integrated}) to fit the models, where the gluon density is obtained by evaluating Eq.~(\ref{eq:dipole gluon}).
%Comparing, the GBW and BGK models, one identifies that, the GBW model is effectively the BGK model with 
%\begin{equation}
%\begin{split}
%    \alpha_s(\mu^2)xg(x,\mu^2)\rightarrow &\alpha_s(Q_0^2)xg(x,Q_0^2)\\
%    &=\alpha_s x^{-\lambda}(1-x)^{5.6}.
%\end{split}
%\end{equation}

Considering the small-$r$ limit of the GBW model, 
\begin{equation}
    \left.\sigma_{\mathrm{GBW}}\left(x,r\right)\right|_{r\ll Q_s}\approx r^2
    Q_s^2/4\,,
\end{equation}
comparison to 
\begin{equation}
\left. \sigma_{\mathrm{dipole}}\left(x,r\right)\right|_{r\ll
Q_s}\approx\frac{r^2 \pi^2\alpha_s(\mu^2)xg(x,\mu^2)}{3}\,,
\end{equation} 
%
from BGK, suggests that the GBW model is an approximation in which
$\alpha_s(\mu^2)$ and $xg(x,\mu^2)$ are independent of
$r$~\cite{Bartels:2002cj}.  
%
For this reason, in one version of the model discussed below, we shall account
for the running coupling in Eq.~(\ref{eq:angle-integrated}), by assuming that
$\alpha_s$ in Eq.~(\ref{eq:dipole gluon}) is constant for the GBW model, and
thus explicitly multiply it by the running coupling
%
\begin{equation}
\alpha_s(\mu^2)\fdp(x,k_T^2)\rightarrow \alpha_s(\mu)\frac{\alpha_s\fdp(x,k_T^2)}{0.2},
\end{equation}
where
%
\begin{equation}
\alpha_s(\mu^2)=\frac{1}{\frac{11 C_A-2n_f}{12\pi}\log\left(\frac{\mu^2}{\Lambda_{\mathrm{QCD}}^2 }\right)},
\end{equation}
and  $\Lambda_{\mathrm{QCD}}^2=0.09\;\GeV^2$.

The factor 0.2 is an arbitrary normalization, whose effect is absorbed by
$\sigma_0$, and hence bears no importance.  (For a more detailed analysis of the
dipole gluon density from the BGK model see
Ref.~\cite{Luszczak:2022fkf}\footnote{In Ref.~\cite{Luszczak:2022fkf}, the
coupling constant was treated differently, and the large $k_T$-region was
matched to the derivative of $xg(x)$, while we have used series transformation
to accelerate the high-$k_T$ integration. Their treatment is also different from
the original BGK paper~\cite{Bartels:2002cj}. As a consequence their gluon
density is positive in the large $k_T$-region and the overall large-$k_T$
behaviour is significantly different.}) While there is some ambiguity on what
the argument of $\alpha_s(\mu^2)$ should be, we follow
Ref.~\cite{Kwiecinski:1997ee} and use
%
\begin{equation}
	\mu^2=k_T^2+{\kappa'}_t^2+m_f^2+\mu^2_0,
\end{equation}
%
where we add $\mu^2_0=4\;\GeV^2$ in order to freeze the coupling at low scales.

@@

We fitted the GBW and BGK models to the $F_2$ data from
HERA~\cite{Abt:2017nkc}. A numerical program was written with help of
CERNLIB~(DPSIPG)~\cite{Kolbig:1972zz}, GSL~\cite{GSL}, CUBA~\cite{Hahn:2004fe}
and ROOT~\cite{Brun:1997pa} libraries to evaluate $F_2$. The fitting was
performed using MnMigrad and MnSimplex of ROOT::Minuit2~\cite{James:2004xla}.

The data were selected to be in the range $0.045\leq Q^2\leq
650\;\mathrm{GeV^2}$, $x<0.01$.  As with the previous fit~\cite{Goda:2022wsc},
the $c$ and $b$ flavours were taken into account with the mass $1.3$ and
$4.6\;\mathrm{GeV}$, respectively. As discussed earlier, we take the light
quarks to be massless.

We studied the following cases
%
\begin{itemize}
\item GBW model with the fixed coupling in $k_T$-factorization  ($k_T$-GBW),
\item GBW model with the running coupling in $k_T$-factorization (rc-$k_T$-GBW),
\item BGK model in $k_T$-factorization ($k_T$-BGK),
\end{itemize} 
%
and, as a reference, we provide the following results from
Ref.~\cite{Goda:2022wsc};
%
\begin{itemize}
\item GBW model with massless light quarks in the dipole factorization ($r$-GBW),
\item GBW model with massive light quarks in the dipole factorization ($r$-GBW-massive),
\item  BGK model in  the dipole factorization ($r$-BGK).
\end{itemize}

%-----------------------------------------------------------------------------
\begin{table}[t]
\begin{subtable}{\textwidth}
\center\footnotesize
\begin{tabular}{|c|c|c|c|c|}
	\hline - & $\sigma_0 $ [mb] & $x_0 \left(10^{-4}\right)$ & $\lambda$ & $\chi^2/\mathrm{dof}$ \\\hline 
	{\footnotesize $r$-GBW} & 1.907e+01& 2.582e+00& 3.219e-01& 4.438e+00\\\hline 
	{\footnotesize $r$-GBW-massive} & 2.384e+01& 1.117e+00& 3.082e-01& 5.274e+00\\\hline 
	{\footnotesize $k_t$-GBW} & 3.344e+01& 1.333e+00& 3.258e-01& 4.396e+00\\\hline 
	{\footnotesize rc-$k_t$-GBW} & 1.520e+01& 2.648e+00& 3.211e-01& 2.447e+00\\\hline 
\end{tabular}
%\caption{}
\vspace{2mm}
\end{subtable}
\begin{subtable}{\textwidth}
\center\footnotesize
\begin{tabular}{|c|c|c|c|c|c|c|}
	\hline - & $\sigma_0 $ [mb] & $A_g$ & $\lambda_g$ & $C$ & $\mu_0^2 \left[\mathrm{GeV^2}\right]$ & $\chi^2/\mathrm{dof}$ \\\hline 
	{\footnotesize $r$-BGK} & 2.326e+01& 1.181e+00& 8.317e-02& 3.294e-01& 1.873e+00& 1.556e+00\\\hline 
	{\footnotesize $k_t$-BGK} & 3.470e+01& 1.048e+00& 2.205e-01&2.391e-01& 9.954e-01&  1.527e+00\\\hline 
\end{tabular}
%\caption{}
\vspace{2mm}
\end{subtable}
\caption{Fit parameters of respective models. The parameters of the
dipole factorization cases are from Ref.~\cite{Goda:2022wsc}.}
\label{tab:table}
\end{table}
%-----------------------------------------------------------------------------
